---
author: "Ryan Christensen"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Load and Tidy the shot data
```{r message=FALSE}
library(tidyverse)
library(plotly)

theme_set(theme_minimal())

shot_data <- readxl::read_excel("Sample Player Test Data.xlsx")

shot_data <- shot_data %>% 
  mutate_at(c("Player", "Club"), factor) 
```

```{r eval=FALSE, echo=FALSE}
glimpse(shot_data)
```

Assuming that the goal of any club test is to find better accuracy, more distance, or some combination of both. I'm choosing to start by looking at accuracy.      

## Average distance offline for each club

```{r}
shot_data %>% 
  select(Player, Club, `Actual Offline`) %>%
  group_by(Club) %>% 
  summarise(Avg_Miss = mean(`Actual Offline`, na.rm = TRUE))
```

It looks like there might be one club that is much closer to the target than the other. Next, I want to see what the distribution of this offline data looks like.    

```{r}
pl <- ggplot(shot_data, aes(`Actual Offline`)) +
  geom_density() +
  geom_vline(xintercept = mean(shot_data$`Actual Offline`), color = 'red', alpha = 0.34) +
  scale_x_continuous(limits = c(-70, 70))

ggplotly(pl)
```

Looks more or less normal which means that I can use a T-Test to verify what the simple means above are hinting at. First I'll check that the distribution is normal using a `Shapiro-Wilk` test.  

```{r}
shapiro.test(shot_data$`Actual Offline`)
```

## T-Test: 

```{r}
t_test <- t.test(shot_data$`Actual Offline`~shot_data$Club)
t_test
```

Cool, so it looks like Club B is up closer to the target on average for most of the participants. I like to see results like this visually so I will now make a couple of visuals to help enforce this result.    

## Boxplot showing the range of the actual offline shot for each club

```{r}
p <- shot_data %>% 
  group_by(Player, Club) %>% 
  summarise(Avg_Miss = mean(`Actual Offline`)) %>% 
  ungroup() %>% 
  ggplot(aes(Club, Avg_Miss, col = Club)) + 
    geom_boxplot() +
    coord_flip()

ggplotly(p)
```

## Plot of the average offline total per each player

```{r warning=FALSE}
offline <- shot_data %>% 
  group_by(Player, Club) %>% 
  summarise(Avg_Miss = mean(`Actual Offline`)) %>% 
  ungroup()

plot <- ggplot(offline, aes(Club, Avg_Miss, group = Player)) +
    geom_point(aes(text = paste("Avg Miss: ", Avg_Miss, ", Player: ", Player), color = Player, size = 3, alpha = 0.5)) +
    geom_line(aes(color = Player, alpha = 0.5)) +
    coord_flip() +
    theme(legend.position = "none")

ggplotly(plot, tooltip = c("text"))
```

Pretty clear that Club B is biasing shots more left than Club A. There could be some left-handed players which might explain some of the players going farther right than they started. I want to see if there was any effect on the top ten players with the most significant accuracy gains as far as distance is concerned.         

```{r}
diff_df <- shot_data %>% 
  group_by(Player, Club) %>% 
  summarise(Avg_Miss = mean(`Actual Offline`)) %>% 
  spread(Club, Avg_Miss) %>% 
  mutate(Diff_Miss = `Club A` - `Club B`) %>% 
  arrange(desc(abs(Diff_Miss)))

dva <- shot_data %>% 
  right_join(diff_df) %>% 
  select(Player, Club, `Actual Carry`, Diff_Miss) %>% 
  group_by(Player, Club, Diff_Miss) %>% 
  summarise(Avg_carry = mean(`Actual Carry`)) %>% 
  spread(Club, Avg_carry) %>% 
  mutate(Diff_Carry = `Club A` - `Club B`) %>% 
  select(Player, Diff_Miss, Diff_Carry) %>% 
  arrange(desc(abs(Diff_Carry)))
```

Interestingly it seems that the players with the biggest accuracy gains had the smallest change in carry yards in the group.     

```{r}
ggplot(dva, aes(Diff_Miss, Diff_Carry)) +
  geom_point(aes(col = Player, size = 3, alpha = 0.34)) + 
  geom_smooth(method = lm, color = "FireBrick") +
  theme(legend.position = 'none')
```

A quick plot shows a general downward trend suggesting that there is a negative correlation in this case between accuracy gained and distance delta. However the confidence interval surrounding the line is massive, so I'm not sure I could make a definitive statement here. What does seem to be clear is that Club B averages about 7 yards closer to the target than Club A.   

```{r eval=FALSE}

shot_data %>% 
  filter(Player %in% c("Player 4", "Player 12", "Player 14", "Player 16")) %>% 
  select(Player, Club, `Ball Speed`, `Vertical Launch`, `Horizontal Launch`, `Actual Carry`, `Spin Rate`, `Spin Axis`) %>% 
  group_by(Player, Club) %>% 
  summarise(Avg_Carry = mean(`Actual Carry`), Avg_HL = mean(`Horizontal Launch`), Avg_SA = mean(`Spin Axis`))
```




